Developing effective theories for single and bilayer graphene using density matrix downfolding

### 2019-09-10 draft.tex
candidate descriptors cleaned, need to just do wf sampling

\vspace{-10pt}
\paragraph{Selecting candidate descriptors}
We selected the following set of 1- and 2-body operators to construct our candidate descriptors based on our understanding of the low-energy excitations of the CuO molecule and our sampled states:
\begin{equation}
\begin{split}
\sum_{\sigma = \uparrow, \downarrow} \Bigg(\epsilon_{3d_\pi}\hat{n}_{3d_\pi,\sigma} + \epsilon_{3d_{z^2}}\hat{n}_{3d_{z^2},\sigma} +  \epsilon_{2p_z} \hat{n}_{2p_z,\sigma} + \epsilon_{2p_\pi}\hat{n}_{2p_\pi,\sigma} + \epsilon_{4s}\hat{n}_{4s,\sigma} +\\
t_\pi \hat{c}_{3d_\pi,\sigma}^\dagger \hat{c}_{2p_\pi,\sigma} + t_{dz} \hat{c}_{3d_{z^2},\sigma}^\dagger \hat{c}_{2p_z,\sigma}  + t_{sz}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{2p_z,\sigma} + t_{sd}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{3d_{z^2},\sigma} + \text{h.c.} + \Bigg)  \\
J_{sd}\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i} + U_s \hat{n}_{4s,\uparrow}\hat{n}_{4s,\downarrow} + E_0.
\end{split}
\label{eq:models}
\end{equation}
The operator subscripts correspond to the intrinsic atomic orbital (IAO) basis elements constructed using the molecular orbitals in \eqref{eq:sampling}, and can be seen in \ref{fig:cuo_regr}a.

The parantheses of \eqref{eq:models} contain the full set of symmetry allowed 1-body operators within the Cu 3d, 4s and O 2p space.
These operators are required to describe the most basic single particle features seen in molecules like orbital occupations and hybridizations.
Two 2-body interaction terms are also considered: a Hund's coupling between Cu 4s and 3d orbitals and a Cu 4s Coulomb repulsion.
The prior term is motivated by the existence of a $\sim $ 0.5 eV Hund's coupling on the neutral copper atom.
Motivation for the latter comes from the FN-DMC energies of our sampled data.
We find that states with a Cu $3d^{10} 4s^{2}$ configuration lie at least 7 eV above our lowest energy sampled state, while the highest energy states with a $3d^{10} 4s^{1}$ configuration lie near 2 eV.

>>>>

\vspace{-10pt}
\paragraph{Selecting candidate descriptors}
The following set of 1- and 2-body operators were selected as candidates for the model based on understanding of the low-energy excitations of the CuO molecule and properties of the sampled low-energy states:
\begin{equation}
\begin{split}
\sum_{\sigma = \uparrow, \downarrow} \Bigg(\epsilon_{3d_\pi}\hat{n}_{3d_\pi,\sigma} + \epsilon_{3d_{z^2}}\hat{n}_{3d_{z^2},\sigma} +  \epsilon_{2p_z} \hat{n}_{2p_z,\sigma} + \epsilon_{2p_\pi}\hat{n}_{2p_\pi,\sigma} + \epsilon_{4s}\hat{n}_{4s,\sigma} +\\
t_\pi \hat{c}_{3d_\pi,\sigma}^\dagger \hat{c}_{2p_\pi,\sigma} + t_{dz} \hat{c}_{3d_{z^2},\sigma}^\dagger \hat{c}_{2p_z,\sigma}  + t_{sz}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{2p_z,\sigma} + t_{sd}\hat{c}_{4s,\sigma}^\dagger \hat{c}_{3d_{z^2},\sigma} + \text{h.c.} + \Bigg)  \\
J_{sd}\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i} + U_s \hat{n}_{4s,\uparrow}\hat{n}_{4s,\downarrow} + E_0.
\end{split}
\label{eq:models}
\end{equation}
The operator subscripts correspond to the intrinsic atomic orbital (IAO) basis elements constructed using the molecular orbitals in \eqref{eq:sampling}, and can be seen in \ref{fig:cuo_regr}a.

The first chunk of descriptors were selected using physical intuition and symmetry principles.
As described above, the low-lying excitations in CuO observed in experiment involve the Cu 3d, 4s and O 2p atomic orbitals.
Given that orbitals also hybridize, the occupation energies of and hopping terms between these orbitals were selected as candidate operators.
Symmetry considerations help narrow down the set as hopping terms are only non-zero between same symmetry orbitals, and occupation energies can be traced over symmetry related orbitals, for example $n_{2p_\pi} = n_{2p_x} + n_{2p_y}$.
The term $\sum_{i\in {\{\text{xy, xz, ...}}\}} \vec{S}_{4s} \cdot \vec{S}_{3d_i}$ was also considered, motivated by the existence of a $\sim $0.5 eV Hund's coupling observed on the copper atom \textbf{ref}.

The inclusion of a double occupancy term $n_{4s,\uparrow} n_{4s,\downarrow}$ was motivated directly from the sampled low-energy states.
Within the sampled data, states with a Cu $4s^{2}$ configuration lie at least 7 eV above the lowest energy sampled state, while the lowest energy state with a $4s^{1}$ configuration lies at 1.6 eV.
If the double occupancy term was not included, the effective theory would predict the prior state to have an energy of $\sim $ 3eV, an underestimation of 4 eV from the \textit{ab-initio} result.
The effective on-site Coulomb repulsion $U_s$ can ensure our model places all $4s^{2}$ states high in energy.

### 2019-09-10 reoutline.tex
candidate descriptors cleaned, need to just do wf sampling

%candidate descriptors 
\item Physical intuition and symmetry

\item Statistics

\item Final set and basis elements

>>>>

%candidate descriptors 
\item The following set of 1- and 2-body operators were selected as candidates for the model based on understanding of the low-energy excitations of the CuO molecule and properties of the sampled low-energy states.

\item The first chunk of descriptors were selected using physical intuition and symmetry principles.

\item The inclusion of a double occupancy term $n_{4s,\uparrow} n_{4s,\downarrow}$ was motivated directly from the sampled low-energy states.
